Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is a merged and final version of two conference papers, \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}, with a unified and streamlined presentation and expanded background materials.}

### 2020-08-24 journal paper/content/related_work.tex
add keller-rady reference and address a couple of comments on intro

The three-way tradeoff between exploration, exploitation and incentives has been studied in several settings other than ours:
incentivizing exploration in a recommendation system
    \citep[\eg][]{Che-13,Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17,Bahar-ec16,Jieming-unbiased18},
dynamic auctions
    \cite[\eg][]{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},
pay-per-click ad auctions with unknown click probabilities
    \cite[\eg][]{MechMAB-ec09,DevanurK09,Transform-ec10-jacm},
coordinating search and matching by self-interested agents
    \citep{Bobby-Glen-ec16},
as well as human computation
    \cite[\eg][]{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
\citet{Bolton-econometrica99,Keller-econometrica05,Johari-ec12} studied models with self-interested agents jointly performing exploration, with no principal to coordinate them.

>>>>

The three-way tradeoff between exploration, exploitation and incentives has been studied in several settings other than ours:
incentivizing exploration in a recommendation system
    \citep[\eg][]{Che-13,Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17,Bahar-ec16,Jieming-unbiased18},
dynamic auctions
    \cite[\eg][]{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},
pay-per-click ad auctions with unknown click probabilities
    \cite[\eg][]{MechMAB-ec09,DevanurK09,Transform-ec10-jacm},
coordinating search and matching by self-interested agents
    \citep{Bobby-Glen-ec16},
as well as human computation
    \cite[\eg][]{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
\citet{Bolton-econometrica99,Keller-econometrica05,Johari-ec12} studied models with self-interested agents jointly performing exploration, with no principal to coordinate them. \gaedit{\cite{keller2003price} studied a model where firms compete via prices where there is a similar exploration vs. competition tension. Besides the difference in competition for quality vs. prices, a crucial difference between our setting and theirs is that in their setting exploration decisions of firms have no bearing on future choices made by consumers.}

### 2020-08-24 journal paper/content/sec-intro.tex
add keller-rady reference and address a couple of comments on intro

We distinguish three classes of bandit algorithms, based on
the prevalent intuition in the area. The distinction concerns the way in which they resolve the fundamental tradeoff between  exploration and \emph{exploitation} (making optimal myopic decisions using the available data). Going from more primitive to more sophisticated, these three classes are as follows:

@@

\subsection{Our Results}
\label{sec:intro-results}

@@

We consider two extremes: a simple Bayesian model with full Bayesian rationality and no performance signals, and a simple frequentist model with reputation scores and no prior knowledge or beliefs. For the theoretical results, the ``no-performance-signals" assumption makes agents' behavior independent of a particular realization of the prior. Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior. Such summarization is essential for formulating lucid and general analytic results, let alone proving them. It is unclear how to incorporate performance signals in a theoretically tractable model. For the numerical results, the \ExptsModel accounts for competition in a more direct way, allows to separate reputation vs. data advantage, and makes our model amenable to numerical simulations. Indeed, simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable.

@@

>>>>

\gaedit{While our model allows for the firms to pick from an arbitrary set of algorithms}, we distinguish between three classes of bandit algorithms, based on the prevalent intuition in the area. The distinction concerns the way in which they resolve the fundamental tradeoff between  exploration and \emph{exploitation} (making optimal myopic decisions using the available data). \gaedit{This classification is important for interpreting the equilibrium algorithm choice of the firms as it allows for a ``low" to ``medium" to ``high" technology interpretation. We further use the classification to pick a representative algorithm from each class for our numerical experiments. Going from low to medium to high, these three classes are as follows:} \gadelete{Going from more primitive to more sophisticated, these three classes are as follows:}

@@

\subsection{Our results}
\label{sec:intro-results}

@@

We consider two extremes: a simple Bayesian model with full Bayesian rationality and no performance signals, and a simple frequentist model with reputation scores and no prior knowledge or beliefs. For the theoretical results, the ``no-performance-signals" assumption makes agents' behavior independent of a particular realization of the prior. Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior. It is important that  Such summarization is essential for formulating lucid and general analytic results, let alone proving them. It is unclear how to incorporate performance signals in a theoretically tractable model. For the numerical results, the \ExptsModel accounts for competition in a more direct way \gaedit{by positing a more realistic model of consumer behavior that does not require them to have direct information on the algorithms deployed by or the bandit problem faced by the firm. It further allows us to separate reputation vs. data advantage and makes our model amenable to numerical simulations.} Indeed, simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable.

@@

\gaedit{Finally, the assumption that firms commit to the algorithms at the start of the game deserves further justification. We impose this commitment assumption since it is more realistic than considering dynamic strategies and algorithms are, by definition, a commitment to a set of rules. In many contexts where such exploration algorithms are typically deployed, consumers arrive at such a rapid volume that it it is infeasible to consider a firm dynamically adapting its algorithm to each consumer that arrives. Indeed, \cite{DS-arxiv} points out how difficult it is to deploy standard exploration algorithms in practice and such dynamic strategies are substantially more difficult to deploy in practice. While it is possible to consider share-adaptive algorithms that would still keep the commitment assumption but allow for indirect dynamic responses to market conditions, such algorithms do not exist currently in the literature but are a good avenue for future work.}
